---
title: "Final Project"
author: "Po-Sheng Lee"
output: github_document
---

```{r package}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, include = FALSE)

library(quantmod)
library(MTS)
library(tidyverse)
library(magrittr)
library(fGarch)
library(fUnitRoots)
library(urca)
library(stargazer)
source("Corner.R")

```

```{r data import}

getSymbols("CPIAUCNS", src = "FRED") # CPI all items
getSymbols("CPILFENS", src = "FRED") # CPI less food and energy
getSymbols("PPIACO", src = "FRED") # PPI all
getSymbols("MCOILWTICO", src = "FRED") # crude oil
getSymbols("XTIMVA01USM664N", src = "FRED") # Import Value Goods for US not adjusted 
getSymbols("IMPCH", src = "FRED") # Import from China
getSymbols("IMP5830", src = "FRED") # Import from Taiwan
getSymbols("IMP5820", src = "FRED") # Import from HK
getSymbols("IMPJP", src = "FRED") # Import from JP
getSymbols("IMPKR", src = "FRED") # Import from Korea

da_cpi <- CPIAUCNS %>%
  as.data.frame() %>%
  rownames_to_column()

da_cpinofood <- CPILFENS %>%
  as.data.frame() %>%
  rownames_to_column()

da_ppi <- PPIACO %>%
  as.data.frame() %>%
  rownames_to_column()

da_oil <- MCOILWTICO %>%
  as.data.frame() %>%
  rownames_to_column()

da_import <- XTIMVA01USM664N %>%
  as.data.frame() %>%
  rownames_to_column()

da_importChina <- IMPCH %>%
  as.data.frame() %>%
  rownames_to_column()

da_importTW <- IMP5830 %>%
  as.data.frame() %>%
  rownames_to_column()

da_importHK <- IMP5820 %>%
  as.data.frame() %>%
  rownames_to_column()

da_importJP <- IMPJP %>%
  as.data.frame() %>%
  rownames_to_column()

da_importKR <- IMPKR %>%
  as.data.frame() %>%
  rownames_to_column()

da <- da_cpi %>%
  left_join(da_cpinofood) %>%
  left_join(da_ppi) %>%
  left_join(da_oil) %>%
  left_join(da_import) %>%
  left_join(da_importChina) %>%
  left_join(da_importTW) %>%
  left_join(da_importHK) %>%
  left_join(da_importJP) %>%
  left_join(da_importKR) %>%
  rename(cpi = CPIAUCNS, cpi_nofood = CPILFENS, ppi = PPIACO, imp_all = XTIMVA01USM664N, oil = MCOILWTICO,
         imp_china = IMPCH, imp_tw = IMP5830, imp_hk = IMP5820, imp_jp = IMPJP, imp_kr = IMPKR) %>%
  mutate(imp_all = imp_all/10^6, imp_china_ratio = imp_china/imp_all) %>%
  column_to_rownames()


```

```{r trade substitute cointergration ECM}

da_substitue <- da %>%
  select(imp_china, imp_tw, imp_hk, imp_jp, imp_kr) %>%
  mutate(imp_china = log10(imp_china), imp_tw = log10(imp_tw), imp_hk = log10(imp_hk), 
         imp_jp = log10(imp_jp), imp_kr = log10(imp_kr)) %>%
  drop_na()

#unit root test

VARorder(da_substitue, maxp = 24) #bic select at order = 13 

t1 <- ur.df(da_substitue[,1], lags = 1, type = "drift")
t2 <- ur.df(da_substitue[,2], lags = 25, type = "drift")
t3 <- ur.df(da_substitue[,3], lags = 15, type = "drift")
t4 <- ur.df(da_substitue[,4], lags = 26, type = "drift")
t5 <- ur.df(da_substitue[,5], lags = 13, type = "drift")
summary(t1)
summary(t2)
summary(t3)
summary(t4)
summary(t5)

coint <- ca.jo(da_substitue, ecdet = "const", type = "trace", K = 2, spec = "transitory") 
# VAR(2) to best describe the co-intergration process
summary(coint)

wt1 <- da_substitue[,1] + 4.02*da_substitue[,2] + 1.95*da_substitue[,3] - 8.04*da_substitue[,4] - 1.31*da_substitue[,5]

wt2 <- da_substitue[,1] - 25.78*da_substitue[,2] + 8.91*da_substitue[,3] + 7.45*da_substitue[,4] + 8.7*da_substitue[,5]

wt3 <- da_substitue[,1] - 10.61*da_substitue[,2] - 0.41*da_substitue[,3] - 4.39*da_substitue[,4] + 1.7*da_substitue[,5]

wt <- cbind(wt1, wt2, wt3)

adfTest(wt1, lags = 3, type = "c")
adfTest(wt2, lags = 3, type = "c")
adfTest(wt3, lags = 3, type = "c")

m1 <- ECMvar1(da_substitue, p = 13, wt)
m2 <- refECMvar1(m1, thres = 1.2)

```


```{r trade substitue China Taiwan Korea}

da_sub_china <- da %>%
  select(imp_all, imp_china, imp_tw, imp_kr) %>%
  mutate(imp_china = imp_china/imp_all, imp_tw = imp_tw/imp_all, imp_kr = imp_kr/imp_all) %>%
  select(-imp_all) %>%
  drop_na()

MTSplot(da_sub_china)

VARorder(da_sub_china, maxp = 24) # select VAR(14)
coint2 <- ca.jo(da_sub_china, ecdet = "const", type = "trace", K = 14, spec = "transitory")
summary(coint2)

m3 <- VAR(diff(as.matrix(da_sub_china)), p = 13)
m4 <- refVAR(m3, thres = 1.2)

Phi_b <- m4$Phi
Sig_b <- m4$Sigma

VARirf(Phi = Phi_b, Sig = Sig_b, lag = 18)


```

```{r trade substitute China seasonal, cache=TRUE}
# seasonal differencing model


m5 <- sVARMA(da_sub_china, order = c(0,1,3), sorder = c(0,1,1), s = 12)
m6 <- refsVARMA(m5, thres = 1.2) # compare with m4


```

```{r trade seasonal east asia, cache=TRUE}

m7 <- sVARMA(da_substitue, order = c(0,1,3), sorder = c(0,1,1), s = 12)
m8 <- refsVARMA(m7, thres = 1.2) # compare with m2

```

```{r inflation and trade}

da_inflation <- da %>%
  select(cpi, imp_all, oil) %>%
  mutate(imp_all = log10(imp_all)) %>%
  drop_na()

VARorder(da_inflation, maxp = 24) # order 17 AIC 3 BIC
adfTest(da_inflation$cpi, lags = 1, type = "c")
adfTest(da_inflation$imp_all, lags = 2, type = "c")
adfTest(da_inflation$oil, lags = 7, type = "c")

ca.jo(da_inflation, ecdet = "const", type = "trace", K = 17, spec = "transitory") %>%
  summary()

#cointergration test is only significant at 0.1

yt <- da_inflation$cpi - 252.29*da_inflation$imp_all + 0.39*da_inflation$oil

m9 <- ECMvar1(da_inflation, p = 17, wt = yt) 
m10 <- refECMvar1(m9, thres = 1.2)

m11 <- VAR(diff(as.matrix(da_inflation)), p = 16) # VAR(17-1) no cointergration only difference model
m12 <- refVAR(m11, thres = 1.2) # aic bic better in this model

Phi_a <- m12$Phi
Sig_a <- m12$Sigma

VARirf(Phi = Phi_a, Sig = Sig_a, lag = 24, orth = FALSE)
VARirf(Phi = Phi_a, Sig = Sig_a, lag = 24)
```

```{r inflation trade seasonal, cache = TRUE}

m13 <- sVARMA(da_inflation, order = c(0,1,3), sorder = c(0,1,1), s = 12)
m14 <- refsVARMA(m13, thres = 1.2) # compare with m12

```


```{r inflation trade china}

da_inflation_china <- da %>%
  select(cpi, imp_china, oil) %>%
  mutate(imp_china = log10(imp_china)) %>%
  drop_na()

VARorder(da_inflation_china, maxp = 24) # order 14

ca.jo(da_inflation_china, ecdet = "const", type = "trace", K = 14, spec = "transitory") %>%
  summary() # second rank of cointergration is borderline case

at1 <- da_inflation_china$cpi + 274.2*da_inflation_china$imp_china - 0.49*da_inflation_china$oil
at2 <- da_inflation_china$cpi - 34.03*da_inflation_china$imp_china + 0.41*da_inflation_china$oil

at <- cbind(at1, at2)

m15 <- ECMvar1(da_inflation_china, p = 14, wt = at) 
m16 <- refECMvar1(m15, thres = 1.2) # this is better 

m17 <- ECMvar1(da_inflation_china, p = 14, wt = at1) 
m18 <- refECMvar1(m17, thres = 1.2) 

m21 <- VAR(diff(as.matrix(da_inflation_china)), p = 13)
m22 <- refVAR(m21, thres = 1.2)

```

```{r seasonal inflation trade china, cache = TRUE}
m19 <- sVARMA(da_inflation_china, order = c(0,1,3), sorder = c(0,1,1), s = 12)
m20 <- refsVARMA(m19, thres = 1.2) # cf m18

```


```{r cpi ppi trade and oil as exogenous}
da_ppi_cpi <- da %>%
  select(cpi, ppi, imp_all, imp_china, oil) %>%
  mutate(imp_china = log10(imp_china), imp_all = log10(imp_all)) %>%
  drop_na() %>%
  select(cpi, ppi) 

da_trade_oil <- da %>%
  select(cpi, ppi, imp_all, imp_china, oil) %>%
  mutate(imp_china = imp_china/imp_all) %>%
  drop_na() %>%
  select(imp_china, oil) 
  

VARXorder(x = da_ppi_cpi, exog = da_trade_oil, maxp = 12, maxm = 3)

m23 <- VARX(zt = da_ppi_cpi, p = 3, xt = da_trade_oil, m = 3)
m24 <- refVARX(m23, thres = 1.2)



```

```{r inflation trade cn ratio}

da_inflation_rchina <- da %>%
  select(cpi, imp_china_ratio, oil) %>%
  drop_na()

MTSplot(da_inflation_rchina)

VARorder(da_inflation_rchina, maxp = 24) # VAR(4) or VAR(17)

ca.jo(da_inflation_rchina, ecdet = "const", type = "trace", K = 17, spec = "transitory") %>%
  summary()

bt <- da_inflation_rchina$cpi - 647.89*da_inflation_rchina$imp_china_ratio - 0.195*da_inflation_rchina$oil

m25 <- ECMvar1(da_inflation_rchina, p = 17, wt = bt)
m26 <- refECMvar1(m25, thres = 1.2) # seems ok
```

```{r inflation trade cn ratio seasonal, cache=TRUE}

m27 <- sVARMA(da_inflation_rchina, order = c(0,1,2), sorder = c(0,1,1), s = 12)
m28 <- refsVARMA(m27, thres = 1.2)

```
# Time Series analysis of trade from East Asia and inflation 

Below is the plot and model of US's import from China, Taiwan HK, JP and Kr. We take the log transformation of the import value. Error Correction model with cointergration series as wess as Seasonal Model are reported here. Seasonal Model with order (0,1,3) , sorder(0, 1, 1) and s = 12 is preferred, for its lower aic and bic.

```{r report asia trade, include= TRUE, cache=TRUE}

MTSplot(da_substitue)

summary(coint)

m2 <- refECMvar1(m1, thres = 1.2)
m8 <- refsVARMA(m7, thres = 1.2)

```

A reduced model with only ratio of China, TW and KR is reported here. No cointergration among these series so a simple VECM and its impulse response function is reported here.


```{r report china tw kr trade ratio, include=TRUE, cache=TRUE}

MTSplot(da_sub_china)

m4 <- refVAR(m3, thres = 1.2)

VARirf(Phi = Phi_b, Sig = Sig_b, lag = 18)

```

CPI, import and oil price is modelled through ECM with cointergration, VEC and seasonal model. VEC is preferred.

```{r report cpi import oil, include=TRUE, cache=TRUE}

MTSplot(da_inflation)

ca.jo(da_inflation, ecdet = "const", type = "trace", K = 17, spec = "transitory") %>%
  summary()

m10 <- refECMvar1(m9, thres = 1.2)
m12 <- refVAR(m11, thres = 1.2) # no cointergration the test value is on borderline
m14 <- refsVARMA(m13, thres = 1.2)

```

CPI, import from china and oil price is modeled. Seasonal and VEC has mixed result. ECM with cointergration is preferred

```{r report cpi import china oil, echo=TRUE, include=TRUE, cache=TRUE}

ca.jo(da_inflation_china, ecdet = "const", type = "trace", K = 14, spec = "transitory") %>%
  summary()

m16 <- refECMvar1(m15, thres = 1.2) # this is better 
m20 <- refsVARMA(m19, thres = 1.2) # for prediction

```

CPI, china trade as ratio, oil. Seasonal Behavior is salient

```{r trade inflation cn ratio, echo=TRUE, include=TRUE, cache=TRUE}

MTSplot(da_inflation_rchina)
m26 <- refECMvar1(m25, thres = 1.2)
m28 <- refsVARMA(m27, thres = 1.2) # this is prefered

```


Use trade and oil as exogenous shock to model the cpi and ppi.

```{r report exogenous shock,echo=TRUE, include=TRUE, cache=TRUE}

MTSplot(da_ppi_cpi)
MTSplot(da_trade_oil)

m24 <- refVARX(m23, thres = 1.2)

```

Prediction.

```{r prediction, echo=TRUE, include=TRUE, cache=TRUE}
sVARMApred(m8, orig = 0, h = 12) # trade among 5 east asian country
sVARMApred(m6, orig = 0, h =12 ) # trade ratio among China, TW and KR
VARpred(m12, h = 12) # CPI imp_all and oil error correction form
sVARMApred(m14, orig = 0, h =12 ) # data the same as m12
sVARMApred(m20, orig = 0, h =12 ) 
sVARMApred(m28, orig = 0, h =12 )


```
